### 2013-12-19 research/paper/results/results-2.3.tex
Committing what I think is my final edit before submission. Kasun needs to check over references

The efficiency of the system can be defined as the ratio of the time average power output to $P_{mean}=\rho U^3D/2$, the kinetic energy in the fluid approaching the body. A similar definition was given in \citet{Barrero-Gil2010a}). The current results show that the system has a peak efficiency of $0.26\%$  for $\reynoldsnumber=165$ and $6.7\%$ for $\reynoldsnumber=22300$. The peak efficiency reported in \citet{Barrero-Gil2010a} for $\reynoldsnumber=22300$ is approximately $5\%$ less than the current result. This difference is due to \cite{Barrero-Gil2010a} using a $3^{rd}$ order polynomial as the interpolating polynomial which under predicts the forcing at values of $\theta$ where the maximum force occurs as compared to the $7^{th}$ order polynomial used in this study.

@@

The natural time scales of the system can be found by solving for the eigenvalues of the linearized equation of motion, namely
\begin{equation}
\label{eqn:eom_linear}
m\ddot{y}{+}c\dot{y}{+}ky{=}\frac{1}{2}\rho U^2 \mathcal{A} a_1\left(\frac{\dot{y}}{U}\right),
\end{equation}
which is simply the equation of motion presented in equation \ref{final_equation_motion} with the polynomial series for the lift force truncated at the linear term, and the forcing term representing vortex shedding removed.

Combining the $\dot{y}$ terms and solving for eigenvalues gives
\begin{equation}
  \label{eqn:eigs}
  \lambda_{1,2}= -\frac{1}{2}\frac{c-\frac{1}{2}\rho U^2\mathcal{A}a_1}{mU}\pm\frac{1}{2}\sqrt{\left[\frac{c-\frac{1}{2}\rho U^2\mathcal{A}a_1}{mU}\right]^2-4\frac{k}{m}}.
\end{equation}

If it is assumed that the spring is relatively weak, $k\rightarrow 0$, a single non-zero eigenvalue remains. This eigenvalue is
\begin{equation}
  \label{eqn:eigs_nospring}
  \lambda=-\frac{c-\frac{1}{2}\rho U^2\mathcal{A}a_1}{mU}.
\end{equation}

Further, if it is assumed that the mechanical damping is significantly weaker than the aerodynamic forces on the body, $c\rightarrow 0$ and
\begin{equation}
  \label{eqn:eigs_nospring_nodamp}
  \lambda=\frac{\frac{1}{2}\rho U^2\mathcal{A}a_1}{mU}.
\end{equation}

@@

All of the quantities that make up $\Gamma_1$ and $\Gamma_2$ can, in theory, be known before an experiments is conducted. However, the quantity $a_1$ is a relatively difficult one to determine, requiring static body experiments or simulations. Here, the geometry is unchanged and results are only being compared at the same \reynoldsnumber. Hence, suitable parameters can be formed by multiplying $\Gamma_1$ and $\Gamma_2$ by $a_1^2$ and $a_1$ respectively, to arrive at a mass-stiffness parameter $\massstiff =  4\pi^2m^{*2}/U^{*2}$, and a mass-damping parameter defined as $\massdamp = c^*m^*$.

@@

Power can be expressed as the product of force and velocity. Therefore the instantaneous power from the fluid to the body can be expressed as $P_t=F_y\dot{y}$. Similarly the dissipated power due to the mechanical damping can be expressed as $P_d=(c\dot{y})\dot{y}$. The time average of these two quantities, described in equations \ref{power_alt} and \ref{power} should be equal due to energy conservation, provided that the mechanical friction losses are neglected. The mean power vs $U^*$ provides a detailed explanation for the variation of the output power when the reduced velocity is increased. The key regions consists of region 1 where the $P_{mean}$ increases with \ustar, region 2 where $P_{mean}$ becomes maximum and region 3 where $P_{mean}$ decreases with \ustar. Time histories of $P_t $ and $P_d$ at each of these key regions are presented in figure \ref{fig:power_time_histories}.

@@

The exact reason behind this discrepancy is unclear. A possible explanation is that galloping is reasonably weak at $\reynoldsnumber=165$.  It was reported by \cite{Barrero-Gil2009} that galloping only starts to occur ar Re $\geq 159$. It may be that simulations conducted so close to the threshold for the onset of galloping are very sensitive to errors due to blockage or under-resolution, both of which can change the effective \reynoldsnumber of the flow. Small changes in effective \reynoldsnumber\ can have a significant impact of the fluid damping, and therefore on the amplitude and transverse velocity of the body. As power is a function of $(\dot{y})^2$ the error between QSS and FSI power is compounded.

@@

In this paper, the power transfer of a square body under aero elastic galloping is analysed by solving the quasi-steady state model equations through numerical integration. At higher $m^*$ ($m^*>30$ at lower \reynoldsnumber and $m^*> 50$ at high \reynoldsnumber) the power output of the system is not dependent on \ustar or natural frequency of the system, but controlled by the non-dimensionalised mass-damping constant \massdamp . By analysing key regions of the power vs \ustar curve is concluded that in order to obtain an optimum power output, the mass-damping constant $\Gamma_2$ should be high, but not so excessive that it hinders the galloping from reaching induced angles of attack where the forcing is significant. The effect of mass ratio was also observed. The peak efficiency was found to be $0.26\%$ for $\reynoldsnumber=165$ and $6.7\%$ for $\reynoldsnumber=22300$  when $\Gamma_2=0.314$ and $\Gamma_2=1.04$ respectively. In the low \reynoldsnumber\ case, the mean power tends to decrease at $m^*<30$ which appears to be due to the influence of vortex shedding. At $\reynoldsnumber=22300$ the opposite is observed where the mean power tends to increase with decreasing mass ratio, as well as the mean power increasing with increasing $U^*$ at low mass ratios. For this higher \reynoldsnumber\ case, When the mass ratio decreases, due to the lower inertia the velocity time trace tend to move from a sinusoidal signal towards a square signal where it sustains high velocities for longer periods of time which leads to a higher mean power output. The limit to peak efficiency was found out to be $13.5\%$ and occurs when $m^*\rightarrow 0$ and $\ustar \rightarrow \infty$ and $\massdamp=1.22$ by analysing the data trend by lowering the $m^*$.

>>>>

The efficiency of the system can be defined as the ratio of the time average power output to $P_{in}=\rho U^3D/2$, the kinetic energy in the fluid approaching the body. A similar definition was given in \citet{Barrero-Gil2010a}). The current results show that the system has a peak efficiency of $0.26\%$  for $\reynoldsnumber=165$ and $6.7\%$ for $\reynoldsnumber=22300$. The peak efficiency reported in \citet{Barrero-Gil2010a} for $\reynoldsnumber=22300$ is approximately $5\%$ less than the current result. This difference is due to \cite{Barrero-Gil2010a} using a $3^{rd}$ order polynomial as the interpolating polynomial which under predicts the forcing at values of $\theta$ where the maximum force occurs as compared to the $7^{th}$ order polynomial used in this study.

@@

The natural time scales of the system can be found by solving for the eigenvalues of the linearized equation of motion, namely
\begin{equation}
\label{eqn:eom_linear}
(m+m_a)\ddot{y}{+}c\dot{y}{+}ky{=}\frac{1}{2}\rho U^2 \mathcal{A} a_1\left(\frac{\dot{y}}{U}\right),
\end{equation}
which is simply the equation of motion presented in equation \ref{final_equation_motion} with the polynomial series for the lift force truncated at the linear term, and the forcing term representing vortex shedding removed.

Combining the $\dot{y}$ terms and solving for eigenvalues gives
\begin{equation}
  \label{eqn:eigs}
  \lambda_{1,2}= -\frac{1}{2}\frac{c-\frac{1}{2}\rho U^2\mathcal{A}a_1}{(m+m_a)U}\pm\frac{1}{2}\sqrt{\left[\frac{c-\frac{1}{2}\rho U^2\mathcal{A}a_1}{(m+m_a)U}\right]^2-4\frac{k}{(m+m_a)}}.
\end{equation}

If it is assumed that the spring is relatively weak, $k\rightarrow 0$, a single non-zero eigenvalue remains. This eigenvalue is
\begin{equation}
  \label{eqn:eigs_nospring}
  \lambda=-\frac{c-\frac{1}{2}\rho U^2\mathcal{A}a_1}{(m+m_a)U}.
\end{equation}

Further, if it is assumed that the mechanical damping is significantly weaker than the aerodynamic forces on the body, $c\rightarrow 0$ and
\begin{equation}
  \label{eqn:eigs_nospring_nodamp}
  \lambda=\frac{\frac{1}{2}\rho U^2\mathcal{A}a_1}{(m+m_a)U}.
\end{equation}

For reasonably heavy bodies, the impact of the added mass can also be neglected, to arrive at a definition of $\lambda$ as
\begin{equation}
  \label{eqn:eigs_nospring_nodamp}
  \lambda=\frac{\frac{1}{2}\rho U^2\mathcal{A}a_1}{mU}.
\end{equation}

@@

All of the quantities that make up $\Gamma_1$ and $\Gamma_2$ can, in theory, be known before an experiment is conducted. However, the quantity $a_1$ is a relatively difficult one to determine, requiring static body experiments or simulations. Here, the geometry is unchanged and results are only being compared at the same \reynoldsnumber. Hence, suitable parameters can be formed by multiplying $\Gamma_1$ and $\Gamma_2$ by $a_1^2$ and $a_1$ respectively, to arrive at a mass-stiffness parameter $\massstiff =  4\pi^2m^{*2}/U^{*2}$, and a mass-damping parameter defined as $\massdamp = c^*m^*$.

@@

Power can be expressed as the product of force and velocity. Therefore the instantaneous power from the fluid to the body can be expressed as $P_t=F_y\dot{y}$. Similarly the dissipated power due to the mechanical damping can be expressed as $P_d=(c\dot{y})\dot{y}$. The time average of these two quantities, described in equations \ref{power} and \ref{power_alt} should be equal due to energy conservation, provided that the mechanical friction losses are neglected. The mean power vs $U^*$ provides a detailed explanation for the variation of the output power when the reduced velocity is increased. The key regions consists of region 1 where the $P_{mean}$ increases with \ustar, region 2 where $P_{mean}$ becomes maximum and region 3 where $P_{mean}$ decreases with \ustar. Time histories of $P_t $ and $P_d$ at each of these key regions are presented in figure \ref{fig:power_time_histories}.

@@

The exact reason behind this discrepancy is unclear. A possible explanation is that galloping is reasonably weak at $\reynoldsnumber=165$.  It was reported by \cite{Barrero-Gil2009} that galloping only starts to occur ar Re $\geq 159$. It may be that simulations conducted so close to the threshold for the onset of galloping are very sensitive to errors due to blockage or under-resolution, both of which can change the effective \reynoldsnumber\ of the flow. Small changes in effective \reynoldsnumber\ can have a significant impact of the fluid damping, and therefore on the amplitude and transverse velocity of the body. As power is a function of $(\dot{y})^2$ the error between QSS and FSI power is compounded.

@@

In this paper, the power transfer of a square body under aero elastic galloping is analysed by solving the quasi-steady state model equations through numerical integration. At higher $m^*$ ($m^*>30$ at lower \reynoldsnumber\ and $m^*> 50$ at high \reynoldsnumber) the power output of the system is not dependent on \ustar or natural frequency of the system, but controlled by the non-dimensionalised mass-damping constant \massdamp . By analysing key regions of the power vs \ustar curve is concluded that in order to obtain an optimum power output, the mass-damping constant $\Gamma_2$ should be high, but not so excessive that it hinders the galloping from reaching induced angles of attack where the forcing is significant. The effect of mass ratio was also observed. The peak efficiency was found to be $0.26\%$ for $\reynoldsnumber=165$ and $6.7\%$ for $\reynoldsnumber=22300$  when $\Gamma_2=0.314$ and $\Gamma_2=1.04$ respectively. In the low \reynoldsnumber\ case, the mean power tends to decrease at $m^*<30$ which appears to be due to the influence of vortex shedding. At $\reynoldsnumber=22300$ the opposite is observed where the mean power tends to increase with decreasing mass ratio, as well as the mean power increasing with increasing $U^*$ at low mass ratios. For this higher \reynoldsnumber\ case, When the mass ratio decreases, due to the lower inertia the velocity time trace tend to move from a sinusoidal signal towards a square signal where it sustains high velocities for longer periods of time which leads to a higher mean power output. The limit to peak efficiency was found out to be $13.5\%$ and occurs when $m^*\rightarrow 0$ and $\ustar \rightarrow \infty$ and $\massdamp=1.22$ by analysing the data trend by lowering the $m^*$.
